---
title: "Semiparametric Estimation of Long-Term Treatment Effects"
author: "Jiafeng Chen and David Ritzwoller"
output: 
  pdf_document:
    citation_package: natbib
    latex_engine: pdflatex
    toc: false
    keep_tex: true
date: "`r format(Sys.time(), '%B %d, %Y')`"
geometry: margin=1in
fontsize: 11pt
bibliography: library.bib
vignette: >
  %\VignetteIndexEntry{vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
library("knitr")
knitr::opts_knit$set(self.contained = FALSE)
knitr::opts_chunk$set(tidy = TRUE, collapse=TRUE, comment = "#>",
                      tidy.opts=list(blank=FALSE, width.cutoff=55))
options(warn=-1)
```

The package `longterm` implements the semiparametric estimators of long-term average treatment effects developed in @chen2021semiparametric. 

# Setup

We observe two data sets. One data set contains measurements of the short-term outcomes of a randomized evaluation of a binary treatment of interest. The other data set contains observational measurements of the joint distribution of short-term and long-term outcomes. We are interested in estimating the average treatment effect of the treatment on a long-term outcome of interest, and are willing to impose the identifying assumptions specified in @chen2021semiparametric. @chen2021semiparametric develop a semiparametric estimator of this parameter in this context with the ``Double/Debiased Machine Learning'' construction of @chernozhukov2018double. 

# Example

We illustrate the use of the package using a data set from @banerjee2015multifaceted. The dataset is included in the package as the list `graduation`. Run `?graduation` for a full description of the dataset. Further details on the contents and construction of this data set is available in the Supplemental Appendix to @chen2021semiparametric. See `?longterm` for full descriptions of the arguments and returned values of `longterm`.

```{r}
library(longterm)

## Assign each observation to `experimental` and `observational` groups randomly
graduation$observe <- as.numeric(runif(nrow(graduation)) <= 0.5)

## Choose Pre-treatment covariates, Short-term Outcomes, and Long-term Outcome
X_vars <- c("ctotal_pcmonth_bsl", "cnonfood_pcmonth_bsl", 
            "cfood_pcmonth_bsl",  "cdurable_pcmonth_bsl",
            "asset_index_bsl",    "asset_prod_index_bsl")
S_vars <- c("ctotal_pcmonth_end", "cnonfood_pcmonth_end", 
            "cfood_pcmonth_end",  "cdurable_pcmonth_end",
            "asset_index_end",   "asset_prod_index_end")
Y_var  <- "ctotal_pcmonth_fup"

## Assume treatment is observed in the observational sample
# Average effect for the experimental population
est_1_1 <- longterm(graduation, S_vars = S_vars, X_vars = X_vars, Y_var = Y_var, 
                    obs = TRUE, estimand = TRUE, type = "glmnet")
# Average effect for the observational population
est_1_0 <- longterm(graduation, S_vars = S_vars, X_vars = X_vars, Y_var = Y_var, 
                    obs = TRUE, estimand = FALSE, type = "glmnet")

## Assume treatment is not observed in the observational sample
# Average effect for the experimental population
est_0_1 <- longterm(graduation, S_vars = S_vars, X_vars = X_vars, Y_var = Y_var, 
                    obs = FALSE, estimand = TRUE, type = "glmnet")
# Average effect for the observational population
est_0_0 <- longterm(graduation, S_vars = S_vars, X_vars = X_vars, Y_var = Y_var, 
                    obs = FALSE, estimand = FALSE, type = "glmnet")

## Display results
kable(data.frame(treatment_observed = c("Yes", "Yes", "No", "No"),
                 subpopulation = c("Observational", "Experimental", 
                                   "Observational", "Experimental"),
                 estimate = c(est_1_1$hat_tau, est_1_0$hat_tau, 
                              est_0_1$hat_tau, est_0_0$hat_tau),
                 standard_error = c(est_1_1$se, est_1_0$se, 
                                    est_0_1$se, est_0_0$se),
                 lower_ci = c(est_1_1$ci[1], est_1_0$ci[1], 
                              est_0_1$ci[1], est_0_0$ci[1]),
                 upper_ci = c(est_1_1$ci[2], est_1_0$ci[2], 
                              est_0_1$ci[2], est_0_0$ci[2])))

```
